---
title: "disaster"
author: "Heewon Yoon"
date: "9/2/2021"
output: html_document
---

goal: combine fema, emdat (with severity) and map county level to congressional based on census data: shows whether there was disaster before election or not in a congressional district in that year (from election 00 to 14)

main datasets of use:
fema
emdat
census
primary date

from US Census website:
Congressional districts for the 108th through 112th sessions were established by the states based on the result of the 2000 Census. Congressional districts for the 113th through 116th sessions were established by the states based on the result of the 2010 Census. Boundaries are effective until January of odd number years (for example, January 2015, January 2017, etc.), unless a state initiative or court ordered redistricting requires a change. All states established new congressional districts in 2011-2012, with the exception of the seven single member states (Alaska, Delaware, Montana, North Dakota, South Dakota, Vermont, and Wyoming)

The congressional district boundaries are the same for the 112th, 111th, 110th and 109th Congress of the United States except for Georgia and Texas which changed between the 109th and 110th Congresses. New maps for the 110th Congress (2007-2009) were created only for the states with congressional district boundary changes.

--> based on these information it seems evident that 109-112 is using the same boundaries except for Georgia and Texas on 110th
--> 109-112 (2000 census), 113-116 (2010 census)
--> with 109-112 everything same except for GA and TX 109 v. 110-112

|       | election year | year | term                    |           | boundaries  | natl code | disaster              |
|-------|---------------|------|-------------------------|-----------|-------------|-----------|-----------------------|
| cd108 | Nov 5 2002    | 2002 | Jan 3 2003 - Jan 3 2005 | 2003-2004 | 2000 census | X         | 2001-01-01 - pri date |
| cd109 | Nov 2 2004    | 2004 | Jan 3 2005 - Jan 3 2007 | 2005-2006 | 2000 census | O         | 2003-01-01 - pri date |
| cd110 | Nov 7 2006    | 2006 | Jan 3 2007 - Jan 3 2009 | 2007-2008 | 2000 census | O         | 2005-01-01 - pri date |
| cd111 | Nov 4 2008    | 2008 | Jan 3 2009 - Jan 3 2011 | 2009-2010 | 2000 census | X         | 2007-01-01 - pri date |
| cd112 | Nov 2 2010    | 2010 | Jan 3 2011 - Jan 3 2013 | 2011-2012 | 2000 census | X         | 2009-01-01 - pri date |
| cd113 | Nov 6 2012    | 2012 | Jan 3 2013 - Jan 3 2015 | 2013-2014 | 2010 census | O         | 2011-01-01 - pri date |
| cd114 | Nov 4 2014    | 2014 | Jan 3 2015 - Jan 3 2017 | 2015-2016 | 2010 census | X         | 2013-01-01 - pri date |
| cd115 | Nov 8 2016    | 2016 | Jan 3 2017 - Jan 3 2019 | 2017-2018 | 2010 census | X         | 2015-01-01 - pri date |
| cd116 | Nov 6 2018    | 2018 | Jan 3 2019 - Jan 3 2021 | 2019-2020 | 2010 census | O         | 2017-01-01 - pri date |

concerns:
- in fema dataset, congressional districts are allocated using GIS. but is this up to date to congressional district changes?

possibility:
assign district using GIS for all?


mapping counties to districts
```{r}
setwd("~/Desktop/Yoon/7. 2021Fall/Dissertation Prosepctus/moneypaper")

library(usdata) # to use abbr function
library(tidyverse)

# fips
fips <- read.csv("state-geocodes-v2016.csv")

fips <- fips %>% rename(Region = Census.Bureau.Region.and.Division.Codes.and.Federal.Information.Processing.System..FIPS..Codes.for.States, 
                        Division = X,
                        FIPS = X.1, 
                        Name = X.2)
fips <- fips[-c(1,2,3,4,5),] # remove unnecessary description columns
fips$FIPS <- as.numeric(fips$FIPS)
fips <- fips %>% filter(FIPS >= 1) # include only 51 states

fips$abb <- state2abbr(fips$Name)

fips <- fips %>% arrange(FIPS)

# congressional district by county (from census)
clean_df <- function(file) {
  x <- read.delim(file, sep=",", header=F) %>% 
        rename(state = V1,
               county = V2,
               district = V3) %>% 
         slice(-1:-2) %>% mutate(state=as.numeric(state))
}

cd109 <- clean_df("cou_cd109_natl.txt") # 51 states, census 00
cd110 <- clean_df("cou_cd110_natl.txt") # 51 states, census 00
cd113 <- clean_df("natl_cocd_delim13.txt") # 43 states, census 10
cd116 <- clean_df("natl_cocd_delim16.txt") # 43 states, census 10
# single member states: Alaska, Delaware, Montana, North Dakota, South Dakota, Vermont, and Wyoming

# what are the differences?
anti_join(cd109, cd110) # shows difference in GA, TX
anti_join(cd113, cd116,  by = c("state", "county", "district")) # difference in DC, NC, VA

# cd109: 108~109
# cd110: 110~112
# cd113: 113~116

join <- function(dat){
 dat = left_join(dat, fips %>% dplyr::select(FIPS, abb), by=c("state"="FIPS"))
 dat$cong <- paste0(dat$abb, dat$district)
 return(dat)
}

cd108 <- join(cd109)
cd108$cd <- "cd108"
cd108$year <- "2002"
cd108$anl.begin <- as.Date("2001-01-01")

cd109 <- join(cd109)
cd109$cd <- "cd109"
cd109$year <- "2004"
cd109$anl.begin <- as.Date("2003-01-01")

cd110 <- join(cd110)
cd110$cd <- "cd110"
cd110$year <- "2006"
cd110$anl.begin <- as.Date("2005-01-01")

cd111 <- cd110
cd111$cd <- "cd111"
cd111$year <- "2008"
cd111$anl.begin <- as.Date("2007-01-01")

cd112 <- cd110 
cd112$cd <- "cd112"
cd112$year <- "2010"
cd112$anl.begin <- as.Date("2009-01-01")

cd113 <- join(cd113)
cd113$cd <- "cd113"
cd113$year <- "2012"
cd113$anl.begin <- as.Date("2011-01-01")

cd114 <- cd113
cd114$cd <- "cd114"
cd114$year <- "2014"
cd114$anl.begin <- as.Date("2013-01-01")

cd115 <- cd113
cd115$cd <- "cd115"
cd115$year <- "2016"
cd115$anl.begin <- as.Date("2015-01-01")

cd116 <- join(cd116)
cd116$cd <- "cd116"
cd116$year <- "2018"
cd116$anl.begin <- as.Date("2017-01-01")

cong <- rbind(cd108, cd109, cd110, cd111, cd112, cd113, cd114, cd115, cd116)
cong$year <- as.integer(cong$year)

# add primary dates for each election year
pri <- read.csv("primarydates.csv") # 1994-2018
pri$pri <- as.Date(pri$date, "%m/%d/%Y")


# add primary dates
cong <- left_join(cong, pri %>% dplyr::select(election_year, state_abbr, pri), 
                  by = c("year" = "election_year", "abb" = "state_abbr"))


cong$anl.begin <- as.Date(cong$anl.begin)
cong$pri <- as.Date(cong$pri)

cong <- cong %>% select(-c(X, state))


# save
write.csv(cong, "cong.csv")

```


constructing disaster data
````{r}
fema <- read.csv("FemaWebDeclarationAreas.csv") 
fema <- fema %>% select(disasterNumber, programTypeCode, programTypeDescription, 
                        stateCode, placeCode, placeName, designatedDate)

# convert to dates
fema$designatedDate <- as.Date(substr(fema$designatedDate, 1, 10))

# change errors
fema$designatedDate <- replace(fema$designatedDate, fema$designatedDate == "2120-08-30", "2012-08-30")
fema$designatedDate <- replace(fema$designatedDate, fema$designatedDate == "2904-09-29", "2012-09-29")
fema$designatedDate <- replace(fema$designatedDate, fema$designatedDate == "7004-06-17", "2004-06-17")
fema$designatedDate <- replace(fema$designatedDate, fema$designatedDate == "8004-06-17", "2004-06-17")
fema$designatedDate <- replace(fema$designatedDate, fema$designatedDate == "9004-09-29", "2004-09-29")

# county code: placecode in fema is 999+countycode
fema$countycode <- as.numeric(str_sub(fema$placeCode, 3))
fema$countycode <- sprintf("%03d", as.numeric(fema$countycode))


# fema_declaration # add this for more info on disasters
fema_dec <- read.csv("FemaWebDisasterDeclarations.csv")
fema_dec <- fema_dec %>% select(disasterNumber, declarationDate, disasterName, 
                                incidentBeginDate, incidentEndDate, declarationType,
                                stateCode, incidentType)

# extract only dates
fema_dec$declarationDate <- as.Date(substr(fema_dec$declarationDate, 1, 10))
fema_dec$incidentBeginDate <- as.Date(substr(fema_dec$incidentBeginDate, 1, 10))
fema_dec$incidentEndDate <- as.Date(substr(fema_dec$incidentEndDate, 1, 10))


# join fema and fema_dec
fema_join <- left_join(fema, fema_dec, by=c("disasterNumber"="disasterNumber"))
# fema is county level, fema_dec is disaster level
# fema: The date a disaster was officially designated (designatedDate)
# fema_dec: The date the disaster was declared (declarationDate)

fema_join <- fema_join %>% filter(declarationType == "Major Disaster") %>%
  select(disasterNumber, programTypeCode, stateCode.x, countycode, placeName, 
         disasterName, designatedDate, declarationDate, incidentBeginDate, incidentEndDate,
         declarationType, incidentType)

fema_join <- distinct(fema_join) # remove duplicates

# after 2000s
fema_join <- fema_join %>% filter(declarationDate >= "2000-01-01")

# emdat for more info
emdat <- read.csv("emdat_public_trial.csv", na.strings=c("", "NA"))
emdat <-  emdat %>% filter(Country == "United States of America (the)") 


library(data.table) # or reshape2
emdat <- melt(emdat, id.vars=c("Dis.No", "Year", "Seq", "Disaster.Group", "Disaster.Subgroup", 
                            "Disaster.Type", "Disaster.Subtype", "Disaster.Subsubtype",
                            "Event.Name",  "Entry.Criteria", "Country", "ISO", "Region", 
                            "Continent", "X", "Origin", "Associated.Dis", "Associated.Dis2", 
                            "OFDA.Response", "Appeal", "Declaration", "Aid.Contribution",
                            "Dis.Mag.Value", "Dis.Mag.Scale", "Latitude", "Longitude", 
                            "Local.Time", "River.Basin", "Start.Year", "Start.Month", 
                            "Start.Day", "End.Year", "End.Month", "End.Day", "Total.Deaths", 
                            "No.Injured", "No.Affected", "No.Homeless", "Total.Affected", 
                            "Reconstruction.Costs...000.US..", "Insured.Damages...000.US..", 
                            "Total.Damages...000.US..", "CPI"))
emdat <- emdat %>% drop_na("value")

library(usdata)
emdat$abb <- state2abbr(emdat$value)
emdat <- emdat %>% select(Year, Seq, Disaster.Subgroup, Disaster.Type, Disaster.Subtype, Event.Name,
                          Entry.Criteria, Declaration, Total.Deaths, No.Injured, 
                          No.Affected, Total.Affected, Total.Damages...000.US.., abb,
                          Start.Year, Start.Month, Start.Day, End.Year, End.Month, End.Day)

# after 2000s
emdat <- emdat %>% filter(Start.Year >= 2000)


## join fema_and emdat_clear

fema_join$year <- as.integer(substr(fema_join$incidentBeginDate, start=1, stop=4))
fema_join$month <- as.integer(substr(fema_join$incidentBeginDate, start=6, stop=7))


# join fema and emdat
disaster <- left_join(fema_join, emdat, by=c("stateCode.x"="abb", 
                                        "year"="Start.Year",
                                        "month"="Start.Month"))

# only include intersections (disasters that exists in both)
disaster <- disaster %>% drop_na("Seq")

disaster$declarationDate <- as.Date(disaster$declarationDate)
disaster$incidentBeginDate <- as.Date(disaster$incidentBeginDate)
disaster$incidentEndDate <- as.Date(disaster$incidentEndDate)

disaster$countycode[disaster$countycode=="0NA"] <- NA

disaster$countycode <- as.integer(disaster$countycode)



disaster <- distinct(disaster)


# operationalize disaster

# total death
# median: 12, mean: 46, max: 1833
disaster$death.med <- ifelse(disaster$Total.Deaths >= 12, 1, 0)
disaster$death.mean <- ifelse(disaster$Total.Deaths >= 46, 1, 0)

# total damages
# median: 1600000, mean: 6904767, max: 125000000
disaster$damage.med <- ifelse(disaster$Total.Damages...000.US.. >=1600000, 1, 0)
disaster$damage.mean <- ifelse(disaster$Total.Damages...000.US.. >= 6904767, 1, 0)

# injured, affected has too many missing values

disaster$duration <- disaster$incidentEndDate - disaster$incidentBeginDate

disaster <- disaster %>% select(stateCode.x, countycode, disasterName, incidentType, 
                                declarationDate, incidentBeginDate,
                                Total.Deaths, Total.Damages...000.US.., 
                                death.med, death.mean, damage.med, damage.mean, duration)

# save
write.csv(disaster, "disaster.csv")

```


assigning disasters to counties

```{r}
disaster <- disaster %>% select(-c(X))
cong <- cong %>% select(-c(X))

# function to assign disaster treatment
get_damage <- function(x) {
  dis <- left_join(cong %>% filter(cd==x), disaster,
                    by=c("abb"="stateCode.x", "county"="countycode"))
  dis <- dis %>% filter(declarationDate > anl.begin &
                              declarationDate < pri)
  dis <- distinct(dis)
  
  dis_a <- dis %>% group_by(cong) %>% 
    summarize(death1 = mean(death.mean),
            death2 = mean(death.med),
            damage1 = mean(damage.mean),
            damage2 = mean(damage.med)) %>%
    mutate(c.death1 = ifelse(death1>0.5, 1, 0),
         c.death2 = ifelse(death2>0.5, 1, 0),
         c.damage1 = ifelse(damage1>0.5, 1, 0),
         c.damage2 = ifelse(damage2>0.5, 1, 0))
  
  cg <- distinct(cong %>% filter(cd==x) %>% select(cd, cong, year, anl.begin, pri))
  d <- left_join(cg, dis_a, by=c("cong"="cong"))
  
  d$c.damage1 <- ifelse(is.na(d$c.damage1)==TRUE, 0, d$c.damage1)
  d$c.damage2 <- ifelse(is.na(d$c.damage2)==TRUE, 0, d$c.damage2)
  d$c.death1 <- ifelse(is.na(d$c.death1)==TRUE, 0, d$c.death1)
  d$c.death2 <- ifelse(is.na(d$c.death2)==TRUE, 0, d$c.death2)

  return(d)
}

d108 <- get_damage("cd108")
d109 <- get_damage("cd109")
d110 <- get_damage("cd110")
d111 <- get_damage("cd111")
d112 <- get_damage("cd112")
d113 <- get_damage("cd113")
d114 <- get_damage("cd114")
d115 <- get_damage("cd115")
d116 <- get_damage("cd116")

dis <- rbind(d108, d109, d110, d111, d112, d113, d114, d115, d116)


# save
write.csv(dis, "dis.csv")


```






initial model of get_damage
```{r}
# cd108
# cong %>% filter(cd=="cd108") %>% distinct(cong) 
# 436 distinct congressional districts

# map counties to districts
dis108 <- left_join(cong %>% filter(cd=="cd108"), disaster, 
                    by=c("abb"="stateCode.x", "county"="countycode"))

# filter only disasters that are within date range
dis108 <- dis108 %>% filter(declarationDate > anl.begin &
                              declarationDate < pri)
dis108 <- distinct(dis108)

# dealing with az01, az02
di108 <- dis108 %>% filter(cong %in% c("AZ01", "AZ02"))
di108 <- aggregate(di108, by=list(county=di108$county, district=di108$district, abb=di108$abb),
                 FUN=min, na.rm=TRUE)
di108 <- di108[-c(1:3)]

dis108 <- dis108 %>% filter(!cong %in% c("AZ01", "AZ02"))
dis108 <- rbind(dis108, di108)

# treatment=1 when more than half the counties within the district experienced disaster
dis108a <- dis108 %>% group_by(cong) %>% 
  summarize(death1 = mean(death.mean),
            death2 = mean(death.med),
            damage1 = mean(damage.mean),
            damage2 = mean(damage.med)) %>%
  mutate(c.death1 = ifelse(death1>0.5, 1, 0),
         c.death2 = ifelse(death2>0.5, 1, 0),
         c.damage1 = ifelse(damage1>0.5, 1, 0),
         c.damage2 = ifelse(damage2>0.5, 1, 0))

# get disaster data for the closest disaster (treatment=1 중 가장 최근)
##### might need to modify later on
# dis108b <- dis108 %>% group_by(cong) %>% arrange(desc(declarationDate)) %>% slice(which.max(death.mean))
# disaster might not be the same for other disaster operationalizations


# then rejoin to cong
cg108 <- distinct(cong %>% filter(cd=="cd108") %>%select(cd, cong, year, anl.begin, pri))
  
d108 <- left_join(cg108, dis108a, by=c("cong"="cong"))

d108$c.damage1 <- ifelse(is.na(d108$c.damage1)==TRUE, 0, d108$c.damage1)
d108$c.damage2 <- ifelse(is.na(d108$c.damage2)==TRUE, 0, d108$c.damage2)
d108$c.death1 <- ifelse(is.na(d108$c.death1)==TRUE, 0, d108$c.death1)
d108$c.death2 <- ifelse(is.na(d108$c.death2)==TRUE, 0, d108$c.death2)


## remove NA entries or not?? 
# problem is NAs are going to be rendered to treatment=0




```





attempt to join disaster to congressional districts
```{r}
library(fuzzyjoin)

df_list <- list()
states <- unique(cong$abb)
for(i in 1:length(unique(cong$abb))) {
print(i)
d <- fuzzy_left_join(cong %>% filter(abb==states[i]), disaster %>% filter(stateCode.x==states[i]),
                by = c("county" = "countycode", 
                       "anl.begin" = "declarationDate",
                       "pri" = "declarationDate"),
                match_fun = list(`==`, `>`, `<=`))
df_list[[i]] <- d
}

binded <- bind_rows(df_list)

# multiple disasters in one district-cycle?

```

