---
title: "interdistrict"
author: "Heewon Yoon"
date: "8/27/2021"
output: html_document
---

Bring from RSQL
```{r}
dime$cycle[dime$bonica_rid=="cand880"]
### subset data using RSQL ####

conn <- dbConnect(RSQLite::SQLite(), "dime.sqlite3")
dbListTables(conn)                  

# get colnames of contribDB
contrib <- dbGetQuery(conn, "SELECT * from contribDB limit 10")
colnames(contrib)

# subset data
dime <- dbGetQuery(conn, "SELECT cycle, transaction_id, amount, date, bonica_cid,
           contributor_type, contributor_gender, contributor_city, contributor_state, 
           contributor_zipcode, recipient_name, bonica_rid, recipient_party, 
           recipient_type, recipient_state, seat, election_type, contributor_district_90s, 
           contributor_district_00s, contributor_district_10s, contributor_cfscore, 
           candidate_cfscore FROM contribDB WHERE contributor_type = 'I' AND recipient_type = 'cand' 
                    AND seat = 'federal:house' AND election_type = 'P'")

# save subsetted data
write.csv(dime, "dime_1014.csv")
```

Data Cleaning
```{r}
setwd("~/Desktop/Yoon/5. 2020Fall/SIG Paper/Analysis")
library(tidyverse)

dime <- read.csv("dime_1014.csv") # 1980-2014
recp <- read.csv("dime_recipients_1979_2014.csv") # 1979-2014

# convert to dates
dime$date <- as.Date(dime$date)

# subset to 1999-01-01 ~ 2014-12-12 (since disaster has data til then)
dime <- subset(dime, cycle >= 2000) # 2000-2014
recp <- recp %>% filter(cycle >= 2000)

# clean recp
recp <- subset(recp, select=c(election, cycle, fecyear, Cand.ID, bonica.rid, 
                              bonica.cid, name, lname, ffname, fname, party, 
                              state, seat, district, Incum.Chall, recipient.cfscore, 
                              cand.gender, total.disbursements, contribs.from.candidate, 
                              unitemized, total.receipts, total.indiv.contrib, 
                              ran.primary, ran.general, p.elec.stat, gen.elec.stat, 
                              gen.elect.pct, recipient.type))
#recp <- subset(recp, seat=="federal:house")
# district is recipient district

# clean dime
dime$d1 <- ifelse(dime$cycle %in% c(2000, 2002, 2004, 2006, 2008), dime$contributor_district_00s, NA)
dime$d2 <- ifelse(dime$cycle %in% c(2010, 2012, 2014), dime$contributor_district_10s, NA)
dime$con_district <- coalesce(dime$d1, dime$d2)
# con_district is contributor district

# merge dime+recp
names(recp)[names(recp) == "bonica.rid"] <- "bonica_rid"
join <- left_join(dime, recp, by=c('cycle', 'bonica_rid'))
join <- join %>% filter(seat.y=="federal:house")
join <- join %>% select(cycle, amount, date, bonica_cid, contributor_type, contributor_gender, 
                        contributor_city, contributor_state, contributor_zipcode, recipient_name,
                        bonica_rid, recipient_party, recipient_type, recipient_state, seat.x,
                        election_type, contributor_cfscore, candidate_cfscore, 
                        con_district, election, fecyear, Cand.ID, 
                        name, lname,fname, party, district, Incum.Chall, recipient.cfscore,
                        cand.gender, total.disbursements, unitemized, total.receipts,
                        total.indiv.contrib, ran.primary, ran.general, p.elec.stat, gen.elec.stat, 
                        gen.elect.pct, recipient.type)

write.csv(join, "donation.csv")

# check if merged right
# seems to be basically the same
score <- subset(join, select=c(candidate_cfscore, recipient.cfscore))
score$r.cf <- round(score$recipient.cfscore, digits=2)
score$same <- ifelse(score$candidate_cfscore == score$r.cf, TRUE, FALSE)
table(score$same)
false <- subset(score, same==FALSE)
false$dif <- false$candidate_cfscore - false$r.cf
table(false$dif) # difference mostly minuscule. but 725 obs difference 0.27
```

combine dime with disaster data (made from mapping.rmd)
```{r}
# recipient and contributor data is merged as donation.csv
donation <- read.csv("donation.csv")
disaster <- read.csv("cd.join.csv")

# maybe i wanted to extract year from primary date data and put into disaster
# pri <- read.csv("primarydates.csv")
# pri$pri.date <- as.Date(pri$date, "%m/%d/%Y")
# pri <- subset(pri, election_year <= 2014)
# disaster$pri.date.x <- as.Date(disaster$pri.date.x)
# # disaster <- left_join(disaster, pri %>% dplyr::select(pri.date, election_year),
# #                       by=c("year"="election_year", "pri.date.x"="pri.date"))

# congress: term: election year (where to look)
# 107: 01-03 : 00 (data between 99-01-01 ~ 00-primary-dd)
# 108: 03-05 : 02 (data between 01-01-01 ~ 02-primary-dd)
# 109: 05-07 : 04  (data between 03-01-01 ~ 04-primary-dd)
# 110: 07-09 : 06  (data between 05-01-01 ~ 06-primary-dd)
# 111: 09-11 : 08  (data between 07-01-01 ~ 08-primary-dd)
# 112: 11-13 : 10 (data between 09-01-01 ~ 10-primary-dd)
# 113: 13-15 : 12 (data between 11-01-01 ~ 12-primary-dd) 
# 114: 15-17 : 14 (data between 13-01-01 ~ 14-primary-dd)

data <- left_join(donation, disaster %>% dplyr::select(cong, year, pri.date.x,
                                                       treat, declarationDate, close), 
                  by=c("district"="cong", "cycle"="year"))

write.csv(data, "data.csv")
```


0. how do natural disasters affect donor composition?
$$
\%.of.within = disaster_{it} + \gamma_i + \eta_t + \epsilon
$$
```{r}
data <- read.csv("data.csv")

# district
data$same <- ifelse(data$district==data$con_district, 1, 0)

cand <- data %>% group_by(district, cycle) %>% dplyr::summarize(count=n(), ind=sum(same==1), out=sum(same==0), ind_cont = sum(amount[same==1]), out_cont = sum(amount[same==0]), disaster = max(treat, na.rm=F))

cand$out.num <- cand$out/(cand$ind+cand$out)
cand$out.amount <- cand$out_cont/(cand$ind_cont+cand$out_cont)

summary(lm(rat_out ~ disaster + factor(district) + factor(cycle), data=cand))

summary(lm(ind ~ disaster + factor(district)+factor(cycle), cand))
summary(lm(ind_cont ~ disaster + factor(district)+factor(cycle), cand))

summary(lm(out ~ disaster + factor(district)+factor(cycle), cand))
summary(lm(out_cont ~ disaster + factor(district)+factor(cycle), cand))


# state
data$same.st <- ifelse(data$contributor_state==data$recipient_state, 1, 0)

cand.st <- data %>% group_by(district, cycle) %>% dplyr::summarize(count=n(), ind=sum(same.st==1), out=sum(same.st==0), ind_cont = sum(amount[same.st==1]), out_cont = sum(amount[same.st==0]), disaster = max(treat, na.rm=F))

cand.st$out.num <- cand.st$out/(cand.st$ind+cand.st$out)
cand.st$out.amount <- cand.st$out_cont/(cand.st$ind_cont+cand.st$out_cont)

summary(lm(out.num ~ disaster + factor(district)+factor(cycle), cand.st))
summary(lm(out.amount ~ disaster + factor(district)+factor(cycle), cand.st))


data %>% group_by(district, cycle, bonica_rid) %>% count()


```


1. district-cycle-party level

The effect of natural disaster on winning ideology

$$
W.Ideology_{itp} = \alpha + \beta_1 Disaster_{it} + \gamma_i + \eta_t + \delta_p + \epsilon_{itp}
$$

district $i$, election cycle $t$, party $p$. $\gamma_i$ is district fixed effect, $\eta_t$ is election cycle fixed effect, $\delta_p$ is party fixed effect

The heterogenous effect of timing

$$
W.Ideology_{itp} = \alpha + \beta_1 Disaster_{it} + \beta_2 Timing_{itp} + \beta_3 Disaster_{it} * Timing_{itp} + \gamma_i + \eta_t + \delta_p + \epsilon_{itp}
$$

2. individual candidate level

How natural disasters affect total contributions (and whether this effect varies by the ideology of candidates)

$$
total.amount = \beta_1 Disaster + \beta_2 Ideology + \beta_3 Disaster * Ideology 
$$

How natural disasters affect contribution from out of district donors (and whether this effect varies by the ideology of candidates)

$$
out.of.district.donation = \gamma_1 Disaster + \gamma_2 Ideology + gamma_3 Disaster * Ideology
$$



Some mapping trials from here
The one I ended up using is in mapping.Rmd file. these are some attempts trying to map counties to congressional districts using shapefiles but did not work out well

mapping trial 1 (using USAboundaries)
```{r}
library(USAboundaries)
library(sf)
library(lwgeom)
library(ggmap)
# match counties to congressional districts: fema
# note. USAboundaries only has cd115 districts
fema <- subset(fema, declare.date >= "1994-01-01") # from 1994
# placecode in fema is 999+countycode
fema$countycode <- as.numeric(str_sub(fema$placeCode, 3))

# map counties to congressional districts
union_sf <- us_counties(resolution = 'high')
union_congress_sf <- us_congressional(resolution = "low")
union_sf <- union_sf %>% dplyr::select(name, geometry, countyfp)
union_congress_sf <- union_congress_sf %>% dplyr::select(geoid, state_name, geometry, cd115fp, state_abbr)
cong_join <- st_join(union_sf, union_congress_sf, largest = TRUE)

# make congressional district code similar to that of dime
cong_join$congress <- paste0(cong_join$state_abbr, cong_join$cd115fp)
cong_join$countyfp <- as.numeric(cong_join$countyfp)

# map counties in fema to congressional districts
fema_map <- left_join(fema, cong_join %>% dplyr::select(countyfp, state_abbr, congress), 
                      by = c("countycode" = "countyfp", "stateCode" = "state_abbr"))
fema_map <- subset(fema_map, select=c(programTypeCode, placeName, 
                                      declare.date, countycode, congress, geometry))
```

mapping trial 2 (using tigris)
```{r}
# statefp to state names
library(usdata)
library(tigris)
fips <- read.csv("state-geocodes-v2016.csv")
fips <- fips %>% rename(Region = Census.Bureau.Region.and.Division.Codes.and.Federal.Information.Processing.System..FIPS..Codes.for.States, 
                        Division = X,
                        FIPS = X.1, 
                        Name = X.2)
fips <- fips[-c(1,2,3,4,5,6,7,14,18,19,25,33,34,44,49,54,55,64),]
fips$abb <- state2abbr(fips$Name)

# 113 = year 2013
cong113 <- congressional_districts(year=2013)
coun113 <- counties(year=2013)

cong113 <- subset(cong113, select=c(STATEFP, CD113FP, GEOID, geometry))
cong113 <- left_join(cong113, fips %>% dplyr::select(FIPS, Name, abb),
                     by=c("STATEFP"="FIPS"))
coun113 <- subset(coun113, select=c(STATEFP, COUNTYFP, GEOID, NAME, geometry))
join113 <- st_join(coun113, cong113, largest=TRUE)

join113$district <- paste0(join113$abb, join113$CD113FP)
join113$cdyear <- 113 
join113$year <- "2013-01-01"

# 112 = year 2011
cong112 <- congressional_districts(year=2011)
coun112 <- counties(year=2011)

cong112 <- subset(cong112, select=c(STATEFP, CD112FP, GEOID, geometry))
cong112 <- left_join(cong112, fips %>% dplyr::select(FIPS, Name, abb),
                     by=c("STATEFP"="FIPS"))
coun112 <- subset(coun112, select=c(STATEFP, COUNTYFP, GEOID, NAME, geometry))
join112 <- st_join(coun112, cong112, largest=TRUE)

join112$district <- paste0(join112$abb, join112$CD112FP)
join112$cdyear <- 112
join112$year <- "2011-01-01"

# COUNTYFP in tigris is same as countycode (placecode-99) in fema
# abb in tigris same as stateCode in fema
# match to fema based on declaration date!

# tigris has later than 2010 (for both)
# usaboundaries has 2016 and 1783-2000 county districts
# ucla has congressional districts
```


Mapping trial 3 (using ucla, census)
```{r}
# use ucla and census data

# ucla
# call district boundaries for the congressional districts from a specific congress
get_congress_map <- function(cong=113) {
  tmp_file <- tempfile()
  tmp_dir  <- tempdir()
  zp <- sprintf("http://cdmaps.polisci.ucla.edu/shp/districts%03i.zip",cong)
  download.file(zp, tmp_file)
  unzip(zipfile = tmp_file, exdir = tmp_dir)
  fpath <- paste(tmp_dir, sprintf("districtShapes/districts%03i.shp",cong), sep = "/")
  st_read(fpath)
}

# counties from census
cnty2010 <- st_read("gz_2010_us_050_00_500k.shp")
cnty2000 <- st_read("co99_d00.shp")
cnty2000 <- cnty2000 %>% st_set_crs("NAD83")
cnty1990 <- st_read("co99_d90.shp")
cnty1990 <- cnty1990 %>% st_set_crs("NAD83")

# districts from ucla
cd113 <- get_congress_map(113) #2013 # nad83
cd113 <- cd113 %>% dplyr::select(STATENAME, ID, DISTRICT, STARTCONG, geometry)

cd112 <- get_congress_map(112) #2011
cd112 <- st_transform(cd112, st_crs(cnty2010))
cd112 <- cd112 %>% dplyr::select(STATENAME, ID, DISTRICT, STARTCONG, geometry)

cd111 <- get_congress_map(111) #2009
cd111 <- st_transform(cd111, st_crs(cnty2010))
cd111 <- cd111 %>% dplyr::select(STATENAME, ID, DISTRICT, STARTCONG, geometry)

cd110 <- get_congress_map(110) #2007
cd110 <- st_transform(cd110, st_crs(cnty2010))
cd110 <- cd110 %>% dplyr::select(STATENAME, ID, DISTRICT, STARTCONG, geometry)

cd109 <- get_congress_map(109) #2005
cd109 <- st_transform(cd109, st_crs(cnty2010))
cd109 <- cd109 %>% dplyr::select(STATENAME, ID, DISTRICT, STARTCONG, geometry)

cd108 <- get_congress_map(108) #2003
cd108 <- st_transform(cd108, st_crs(cnty2010))
cd108 <- cd108 %>% dplyr::select(STATENAME, ID, DISTRICT, STARTCONG, geometry)

cd107 <- get_congress_map(107) #2001
cd107 <- st_transform(cd107, st_crs(cnty2010))
cd107 <- cd107 %>% dplyr::select(STATENAME, ID, DISTRICT, STARTCONG, geometry)

cd106 <- get_congress_map(106) #1999
cd106 <- st_transform(cd106, st_crs(cnty2010))
cd106 <- cd106 %>% dplyr::select(STATENAME, ID, DISTRICT, STARTCONG, geometry)

cd105 <- get_congress_map(105) #1997
cd105 <- st_transform(cd105, st_crs(cnty2010))
cd105 <- cd105 %>% dplyr::select(STATENAME, ID, DISTRICT, STARTCONG, geometry)

cd104 <- get_congress_map(104) #1995
cd104 <- st_transform(cd104, st_crs(cnty2010))
cd104 <- cd104 %>% dplyr::select(STATENAME, ID, DISTRICT, STARTCONG, geometry)

cd103 <- get_congress_map(103) #1993 # nad83
cd103 <- cd103 %>% dplyr::select(STATENAME, ID, DISTRICT, STARTCONG, geometry)


# map county to district 
join113 <- st_join(cnty2010, st_make_valid(cd113), largest=TRUE)
join113$abb <- state2abbr(join113$STATENAME)
join113$dis <- sprintf("%02d", as.numeric(join113$DISTRICT))
join113$con_district <- paste0(join113$abb, join113$dis)
join113$year <- "2013-01-01"
join113 <- subset(join113, select=c(COUNTY, NAME, con_district, year))

join112 <- st_join(cnty2010, st_make_valid(cd112), largest=TRUE)
join112$abb <- state2abbr(join112$STATENAME)
join112$dis <- sprintf("%02d", as.numeric(join112$DISTRICT))
join112$con_district <- paste0(join112$abb, join112$dis)
join112$year <- "2011-01-01"
join112 <- subset(join112, select=c(COUNTY, NAME, con_district, year))

join111 <- st_join(cnty2000, st_make_valid(cd111), largest = TRUE) 
join111$abb <- state2abbr(join111$STATENAME)
join111$dis <- sprintf("%02d", as.numeric(join111$DISTRICT))
join111$con_district <- paste0(join111$abb, join111$dis)
join111$year <- "2009-01-01"
join111 <- subset(join111, select=c(COUNTY, NAME,  con_district, year))

join110 <- st_join(cnty2000, st_make_valid(cd110), largest = TRUE) 
join110$abb <- state2abbr(join110$STATENAME)
join110$dis <- sprintf("%02d", as.numeric(join110$DISTRICT))
join110$con_district <- paste0(join110$abb, join110$dis)
join110$year <- "2007-01-01"
join110 <- subset(join110, select=c(COUNTY, NAME, con_district, year))

join109 <- st_join(cnty2000, st_make_valid(cd109), largest = TRUE) 
join109$abb <- state2abbr(join109$STATENAME)
join109$dis <- sprintf("%02d", as.numeric(join109$DISTRICT))
join109$con_district <- paste0(join109$abb, join109$dis)
join109$year <- "2005-01-01"
join109 <- subset(join109, select=c(COUNTY, NAME, con_district, year))

join108 <- st_join(cnty2000, st_make_valid(cd108), largest = TRUE) 
join108$abb <- state2abbr(join108$STATENAME)
join108$dis <- sprintf("%02d", as.numeric(join108$DISTRICT))
join108$con_district <- paste0(join108$abb, join108$dis)
join108$year <- "2003-01-01"
join108 <- subset(join108, select=c(COUNTY, NAME, con_district, year))

join107 <- st_join(cnty2000, st_make_valid(cd107), largest = TRUE) 
join107$abb <- state2abbr(join107$STATENAME)
join107$dis <- sprintf("%02d", as.numeric(join107$DISTRICT))
join107$con_district <- paste0(join107$abb, join107$dis)
join107$year <- "2001-01-01"
join107 <- subset(join107, select=c(COUNTY, NAME, con_district, year))

join106 <- st_join(cnty1990, st_make_valid(cd106), largest = TRUE) 
join106$abb <- state2abbr(join106$STATENAME)
join106$dis <- sprintf("%02d", as.numeric(join106$DISTRICT))
join106$con_district <- paste0(join106$abb, join106$dis)
join106$year <- "1999-01-01"
join106 <- subset(join106, select=c(CO, NAME, con_district, year))
join106$COUNTY <- join106$CO
join106 = select(join106, -1)

join105 <- st_join(cnty1990, st_make_valid(cd105), largest = TRUE) 
join105$abb <- state2abbr(join105$STATENAME)
join105$dis <- sprintf("%02d", as.numeric(join105$DISTRICT))
join105$con_district <- paste0(join105$abb, join105$dis)
join105$year <- "1997-01-01"
join105 <- subset(join105, select=c(CO, NAME, con_district, year))
join105$COUNTY <- join105$CO
join105 = select(join105, -1)

join104 <- st_join(cnty1990, st_make_valid(cd104), largest = TRUE) 
join104$abb <- state2abbr(join104$STATENAME)
join104$dis <- sprintf("%02d", as.numeric(join104$DISTRICT))
join104$con_district <- paste0(join104$abb, join104$dis)
join104$year <- "1995-01-01"
join104 <- subset(join104, select=c(CO, NAME, con_district, year))
join104$COUNTY <- join104$CO
join104 = select(join104, -1)

join103 <- st_join(cnty1990, st_make_valid(cd103), largest = TRUE) 
join103$abb <- state2abbr(join103$STATENAME)
join103$dis <- sprintf("%02d", as.numeric(join103$DISTRICT))
join103$con_district <- paste0(join103$abb, join103$dis)
join103$year <- "1993-01-01"
join103 <- subset(join103, select=c(CO, NAME, con_district, year))
join103$COUNTY <- join103$CO
join103 = select(join103, -1)

cnty.district <- rbind(join103, join104, join105, join106, join107, join108, join109, 
                       join110, join111, join112, join113)

# save as
st_write(cnty.district, "cnty.district.shp", driver="ESRI Shapefile")
```

Mapping counties to districts in fema (by congressional district years)
```{r}
cnty.district <- st_read("cnty.district.shp")
cnty.district$abbr <- substr(cnty.district$cn_dstr, 1, 2)
cnty.district <- cnty.district %>% as_tibble(.) %>% select(-geometry)

# 103
fema$cnty.district[fema$declare.date >= "1993-01-01" & 
                     fema$declare.date < "1995-01-01"] <- "1993-01-01"
# 104
fema$cnty.district[fema$declare.date >= "1995-01-01" & 
                     fema$declare.date < "1997-01-01"] <- "1995-01-01"
# 105
fema$cnty.district[fema$declare.date >= "1997-01-01" & 
                     fema$declare.date < "1999-01-01"] <- "1997-01-01"
# 106
fema$cnty.district[fema$declare.date >= "1999-01-01" & 
                     fema$declare.date < "2001-01-01"] <- "1999-01-01"
# 107
fema$cnty.district[fema$declare.date >= "2001-01-01" & 
                     fema$declare.date < "2003-01-01"] <- "2001-01-01"
# 108
fema$cnty.district[fema$declare.date >= "2003-01-01" & 
                     fema$declare.date < "2005-01-01"] <- "2003-01-01"
# 109
fema$cnty.district[fema$declare.date >= "2005-01-01" & 
                     fema$declare.date < "2007-01-01"] <- "2005-01-01"
# 110
fema$cnty.district[fema$declare.date >= "2007-01-01" & 
                     fema$declare.date < "2009-01-01"] <- "2007-01-01"
# 111
fema$cnty.district[fema$declare.date >= "2009-01-01" & 
                     fema$declare.date < "2011-01-01"] <- "2009-01-01"
# 112
fema$cnty.district[fema$declare.date >= "2011-01-01" & 
                     fema$declare.date < "2013-01-01"] <- "2011-01-01"
# 113
fema$cnty.district[fema$declare.date >= "2013-01-01" & 
                     fema$declare.date < "2015-01-01"] <- "2013-01-01"

# map counties in fema to congressional districts
fema <- subset(fema, declare.date < as.Date("2016-01-01"))
fema$countycode <- as.character(fema$countycode)
fema_map <- left_join(fema, cnty.district %>% dplyr::select(NAME, cn_dstr, year, COUNTY, abbr), 
                      by = c("countycode" = "COUNTY", "stateCode" = "abbr",
                             "cnty.district" = "year"))

fema_map <- subset(fema_map, select=c(programTypeCode, placeName, 
                                      declare.date, countycode, congress, geometry))


```


